---
title: "SNAPPS Script (Cleaning & scoring)"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The aim is to create a script that cleans and scores the SNAPPS survey data downloaded from Bristol Online Survey (BOS). 


#### Load data & packages

```{r message=FALSE, warning=FALSE}

snapps <- read.csv("/Users/mohamedyusuf/R/snapps-test.csv")
library(tidyverse) 
library(knitr)
library(magrittr)
library(kableExtra)

options(knitr.table.format = "html") 

```

#### Check data

```{r }

dim(snapps) # dimensions (rows:columns)
str(snapps) # data structure 

# What the uncleaned data looks like. 

kable(snapps, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = F))

```


Online surveys doesn't allow numbering of the questions, also it has made each input for Q12 into a variable. We will need to clean this up and ensure that it's presentable!


### Data cleaning

##### Change classes of cols

```{r }
# convert Q2 - Q28 & Q30-Q31 into factor 
snapps <- mutate_at(snapps, vars("Q3","Q4","Q5","Q6","Q7","Q8","Q9","Q10","Q11","Q12","Q13","Q14","Q15","Q16","Q17","Q18","Q19","Q20","Q21","Q22","Q23","Q24","Q25","Q26","Q27","Q28","Q30","Q31"), as.factor) 

 # convert Q29_1 - Q29_8 into numeric
snapps <- mutate_at(snapps, vars("Q29_1","Q29_2","Q29_3","Q29_4","Q29_5","Q29_6","Q29_7","Q29_8"), as.numeric)

```


##### Variable sorting

```{r }

# combine scores for the image question (Q12)
snapps$Q12a <- snapps$Q29_1 + snapps$Q29_2 + snapps$Q29_3 + snapps$Q29_5 + snapps$Q29_6 + snapps$Q29_7 #

# drop uneccesarry columns 
snapps <- within(snapps, rm(Q1, Q29_1, Q29_2, Q29_3, Q29_4,Q29_5, Q29_6, Q29_7, Q29_8))

# reorder the columns
snapps <- snapps[,c("Q2","Q3","Q4","Q5","Q6","Q7","Q8","Q9","Q10","Q11","Q12","Q13","Q14","Q15","Q16","Q17","Q18","Q19","Q20","Q21","Q22","Q23","Q24","Q25","Q26","Q27","Q28","Q12a","Q30","Q31")]

# rename columns
colnames(snapps) <- c("Sex","Age","Q2","Q3","Q4","Q5","Q6","Q7","Q8","Q9a","Q9b","Q9c","Q9d","Q9e","Q9f","Q9g","Q9h","Q9i","Q9j","Q9k","Q9l","Q9m","Q9n","Q10a","Q10b","Q11a","Q11b","Q12a","Q12b","Q13")

str(snapps)
# Label the sex variable

snapps$Sex <- factor(snapps$Sex, levels = c("1", "2"), 
                        labels = c("Male", "Female"))

```



### Scoring

```{r , message=FALSE, warning=FALSE}

# Create score column 
snapps$Score <- 0


# Scoring for Q4. If answered 3 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q4[i] == 3){
  snapps$Score[i] <- snapps$Score[i] + 1
  } else  { snapps$Score[i] <- snapps$Score[i] + 0 }
}  

# Scoring for Q5. If answered 1 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q5[i] == 1){
    snapps$Score[i] <- snapps$Score[i] + 1
  } else  { snapps$Score[i] <- snapps$Score[i] + 0 }
}  

# Scoring for Q6. If answered 1 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q6[i] == 1){
    snapps$Score[i] <- snapps$Score[i] + 1
  } else  { snapps$Score[i] <- snapps$Score[i] + 0 }
}  

# Scoring for Q7. If answered 1 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q7[i] == 1){
    snapps$Score[i] <- snapps$Score[i] + 1
  } else  { snapps$Score[i] <- snapps$Score[i] + 0 }
}   

# Scoring for Q8. If answered 1,2,3 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q8[i] == 1){
    snapps$Score[i] <- snapps$Score[i] + 0
  } else  { snapps$Score[i] <- snapps$Score[i] + 1 }
}  

# Scoring for Q10a & Q11a. If answered 1 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q10a[i] == 1 | snapps$Q11a[i] == 1){
    snapps$Score[i] <- snapps$Score[i] + 1
  } else  { snapps$Score[i] <- snapps$Score[i] + 0 }
}  

# Scoring for Q10b & Q11b. If answered 2 then point = 1, else point = 0.
for(i in 1:nrow(snapps)){
  if(snapps$Q10b[i] == 2 | snapps$Q11b[i] == 2){
    snapps$Score[i] <- snapps$Score[i] + 1
  } else  { snapps$Score[i] <- snapps$Score[i] + 0 }
}  

# Scoring for Q12a. Scores are added for this 
for(i in 1:nrow(snapps)){
  snapps$Score[i] <- snapps$Score[i] + snapps$Q12a[i]}


# Make a PFP column showing presence of PFP
snapps$PFP <- 0
for(i in 1:nrow(snapps)){
  if(snapps$Score[i] >= 6) {
    snapps$PFP[i] <- "Yes"
  } else  snapps$PFP[i] <- "No" }  



```


### Cleaned data with Scores

```{r echo=FALSE, message=FALSE, warning=FALSE}

kable(snapps, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = F))


# Save clean data 
write.csv(snapps, file = "snapps.csv", row.names=FALSE)


```


 
 
 
 
 
 
 
 
 
 
 
 
 
 
 

